---
title: "Bird_Project"
author: "Alicia Pentico"
date: "April 21, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Remember that diet is small proxy. Some birds could eat more than others
Moran's I to tell if birds are clustered. 
Diversity analysis?
Add jackknife Analysis
Spatial regression model could have seasonality as factors.

```{r}
library(rgdal) # Provides links to the 'Geospatial' Data Abstraction Library ('GDAL') and access to projection/transformation operations from the 'PROJ.4' library.
library(sp) # Classes and methods for spatial data
library(spdep) # for modeling of spatial dependencies
library(maptools) # provides various mapping functions
library(RANN) # for nearest neighbor calculations
library(gstat)
library(base)
library(leaflet)
library(raster)
```

### READING IN DATA
```{r}
captureBirds<-read.csv("balbina_understory_birds_captures.csv", header = TRUE)
birdInfo<-read.csv("balbina_understory_birds_taxonomy_traits.csv", header = TRUE)
environment<-read.csv("balbina_environmental_variables.csv", header = TRUE)
```

###DATA FORMATTING
```{r}
noRecaptures <- subset(captureBirds, captureBirds$new.individual == "yes")
perLocation<- as.data.frame(table(noRecaptures$site))
perLocation$longitude<- environment$longitude.WGS84
perLocation$latitude<- environment$latitude.WGS84
perLocation$coverage<- environment$forest.cover.200
names(perLocation)<- c("site", "abundance", "longitude", "latitude","coverage")
str(perLocation)
# create vectors to hold species richness and names to later add as cols to perLocation
speciesRichnessBySite=vector('integer',length=length(perLocation$site))
speciesNamesBySite=vector('list',length=length(perLocation$site))
# get number of species at each site, list of species at each site
for (i in 1:length(perLocation$site)){
  siteName = perLocation$site[i]
  
  # operations to get T/F list of species at site
  set<-subset(noRecaptures, site == siteName)
  byLocation <- as.data.frame(table(set$species))
  speciesFoundatSite = (byLocation$Freq != 0)
  
  # add speciesRichness to vector
  speciesRichnessBySite[i] = sum(speciesFoundatSite)
  
  speciesFreqs <- list()
  for (j in 1:length(byLocation$Var1)){
    if(byLocation$Freq[j] != 0){
      speciesFreqs[[as.character(byLocation$Var1[j])]] = byLocation$Freq[j]
    }
  }
  speciesNamesBySite[[i]] = speciesFreqs
  
}
# add species richness and names to perLocation
perLocation$speciesRichness<- speciesRichnessBySite
perLocation$speciesNamesL<- speciesNamesBySite
```



###IDW ABUNDANCE, SPECIES RICHNESS

Creating a spatial points data frame
```{r}
### Converting data into a spatial points data frame
perLocation.spdf <- perLocation
coordinates(perLocation.spdf) <- ~ longitude + latitude ##(remember that in R longitude is assigned to x-axis and latitude to y-axis)
class(perLocation.spdf)
bbox(perLocation.spdf)
```

#Data Transformation Function
```{r}
### Make QQ plots for different transformations
examineNormality = function(data) {
  hist(data)
  qqnorm(data)
  qqline(data)
  
  hist(log(data),main='log')
  qqnorm(log(data), main='log')
  qqline(log(data))
  
  hist(sqrt(data),main='sqrt')
  qqnorm(sqrt(data), main='sqrt')
  qqline(sqrt(data))
  
  hist(1/data,main='1/y')
  qqnorm(1/data, main='1/y')
  qqline(1/data)
}

```

# IDW Variables Setup
```{r}
# set range for interpolation area
minLong = min(perLocation.spdf@coords[,1])
maxLong = max(perLocation.spdf@coords[,1])
minLat = min(perLocation.spdf@coords[,2])
maxLat = max(perLocation.spdf@coords[,2])

x.range <- as.numeric(c(minLong, maxLong))  # min/max longitude of the interpolation area
y.range <- as.numeric(c(minLat, maxLat))  # min/max latitude of the interpolation area

# make grid for idw
grd <- expand.grid(x = seq(from = x.range[1], to = x.range[2], by = 0.001), y = seq(from = y.range[1], 
    to = y.range[2], by = 0.001))  # expand points to grid
coordinates(grd) <- ~x + y
gridded(grd) <- TRUE
plot(grd, cex = 1.5, col = "grey") 
plot(perLocation.spdf, add=TRUE, pch = 1, col = "red", cex = 1)

```

# IDW plot function
```{r}
plotIDW = function(desiredFormula, allDataSPDF, gridToInterpolate, idpRange, titleString) {
  for (idpVal in idpRange) {
    idw <- gstat::idw(formula = desiredFormula, locations = allDataSPDF, newdata = gridToInterpolate, idp=idpVal)
    plot(idw, main=paste(titleString,' for IDP=',idpVal,sep=''))
  }
}
```


```{r}
jackknifeDiffIDP = function(allData, yVarData, yVarName, desiredFormula, idpRange) {
  predValsL = performJackknifeDiffIDP(allData, yVarData, yVarName, desiredFormula, idpRange)
  plotObsVPred(predValsL, yVarData, yVarName, idpRange)
}

performJackknifeDiffIDP = function(allData, yVarData, yVarName, desiredFormula, idpRange) {
  dataLen = length(allData)
  # create a list to hold the lists of predicted vals
  predValsL = replicate(length(idpRange),list())
  # loop over all idp vals
  for (j in 1:length(idpRange)) {
    # get the list of all leave-one-out predicted vals for a particular IDP val
    currIDP = idpRange[j]
    IDW.pred = vector(length = dataLen)
    for (i in 1:dataLen) {
      IDW.pred[i] <- idw(desiredFormula, allData[-i,], allData[i,], idp=currIDP)$var1.pred
    }
    predValsL[[j]] = IDW.pred
  }
  return(predValsL)
}

plotObsVPred = function(predValsL, yVarData, yVarName, idpRange) {
  idpLen = length(idpRange)
  par(mfrow = c(1,idpLen))
  
  for (j in 1:length(idpRange)) {
    currIDP = idpRange[j]
    currPredVals = predValsL[[j]]
    OP = par(pty="s", mar=c(4,3,2,0))
    plot(currPredVals ~ yVarData, asp=1, xlab="Observed", ylab="Predicted", pch=16,
       col=rgb(0,0,0,0.5), main=as.character(paste('IDP=',currIDP,sep='')))
    abline(lm(currPredVals ~ yVarData), col="red", lw=2,lty=2)
    abline(0,1)
  }
  mtext(as.character(paste(yVarName,"IDW Observed vs. Predicted",sep=' ')), outer=TRUE, line=-6)
}
```
#Abundance

```{r}
#Transformation
#examineNormality(perLocation.spdf@data$abundance) # <-- chose sqrt transformation as best
perLocation.spdf@data$transformedAbundance <- sqrt(perLocation.spdf@data$abundance)

#IDW plots
idpRange = seq(2,4,1)
plotIDW(transformedAbundance~1, perLocation.spdf, grd, idpRange, 'IDW plot of sqrt(abundance)')
idw_abundance_2 <- gstat::idw(formula = transformedAbundance~1, locations = perLocation.spdf, newdata = grd, idp=2)

#Jackknife analysis
jackknifeDiffIDP(perLocation.spdf, perLocation.spdf@data$transformedAbundance, 'sqrt(sbundance)', transformedAbundance~1, idpRange)
```

#Species Richness 

```{r}
#Transformation
#examineNormality(perLocation.spdf@data$speciesRichness)
perLocation.spdf@data$transformedRichness <- sqrt(perLocation.spdf@data$speciesRichness)

#IDW plots
idpRange = seq(2,4,1)
plotIDW(transformedRichness~1, perLocation.spdf, grd, idpRange, 'IDW plot of sqrt(richness)')

#Jackknife analysis
jackknifeDiffIDP(perLocation.spdf, perLocation.spdf@data$transformedRichness, 'sqrt(richness)', transformedRichness~1, idpRange)
```

###DIET ANALYSES

Data Wrangling
```{r}
dietL <- c("diet.inv", "diet.vend", "diet.vect", "diet.vfish","diet.vunk","diet.fruit","diet.nect","diet.seed","diet.planto")
dietByRegion <- matrix(1:342, nrow = 38, ncol = 9)
colnames(dietByRegion) <- dietL
for (i in 1:38){ #num rows is each location
  for (j in 1:9){ #num cols is each diet type
    
    food = birdInfo[,dietL[j]] #Vector of particular Diet
    
    sumDiet <- 0
    individualCount <- 0
    
    birdsAtSite <- perLocation$speciesNamesL[[i]]
    birdNames = names(birdsAtSite)
    
    for (k in 1:length(birdNames)){
      bird = birdNames[k]
      numBirds = birdsAtSite[[bird]]
      
      individualCount <- individualCount+numBirds
      
      birdLocation = match(bird, birdInfo$birdlife.2016)
      if (is.na(birdLocation)){
        birdLocation = match(bird, birdInfo$cbro.2015)
      }
      else if (is.na(birdLocation)){
        birdLocation = match(bird, birdInfo$wilman.2014)
      }
      foodEaten = food[birdLocation] #Perent of Diet for that particular bird
      sumDiet <- sumDiet + (numBirds * foodEaten)
      
    }
    
    dietByRegion[i,j] = sumDiet/ (100*individualCount)
  }
}
```

Separating by Diet Type
```{r}
diet.inv <- dietByRegion[,1]
diet.vend <- dietByRegion[,2]
diet.vect <- dietByRegion[,3]
diet.vfishd <- dietByRegion[,4]
diet.vunk <- dietByRegion[,5]
diet.fruit <- dietByRegion[,6]
diet.nect <- dietByRegion[,7]
diet.seed <- dietByRegion[,8]
diet.planto <- dietByRegion[,9]
```

#DIET.INV
```{r}
#Transformation
#examineNormality(diet.inv)
perLocation.spdf@data$diet.inv <- diet.inv ## UNTRANSFORMED DIET IS BEST FOR DIET.INV

#IDW plots
idpRange = seq(2,4,1)
plotIDW(diet.inv~1, perLocation.spdf, grd, idpRange, 'IDW plot of diet.inv')

#Jackknife analysis
jackknifeDiffIDP(perLocation.spdf, perLocation.spdf@data$diet.inv, 'diet.inv', diet.inv~1, idpRange)
```


#DIET.VEND 
TRANSFORMATIONS
```{r}
#Transformation
#examineNormality(diet.vend)
perLocation.spdf@data$diet.vend <- diet.vend ## UNTRANSFORMED DIET IS BEST FOR DIET.VEND

#IDW plots
idpRange = seq(2,4,1)
plotIDW(diet.vend~1, perLocation.spdf, grd, idpRange, 'IDW plot of diet.vend')

#Jackknife analysis
jackknifeDiffIDP(perLocation.spdf, perLocation.spdf@data$diet.vend, 'diet.vend', diet.vend~1, idpRange)
```


#DIET.VECT 
TRANSFORMATIONS
```{r}
#Transformation
#examineNormality(diet.vect)
perLocation.spdf@data$diet.vect <- diet.vect

#IDW plots
idpRange = seq(2,4,1)
plotIDW(diet.vect~1, perLocation.spdf, grd, idpRange, 'IDW plot of diet.vect')

#Jackknife analysis
jackknifeDiffIDP(perLocation.spdf, perLocation.spdf@data$diet.vect, 'diet.vect', diet.vect~1, idpRange)
```


#DET.VFISHD 
TRANSORMATIONS
```{r}
#Transformation
#examineNormality(diet.vfishd)
perLocation.spdf@data$diet.vfishd <- diet.vfishd

#IDW plots
idpRange = seq(2,4,1)
plotIDW(diet.vfishd~1, perLocation.spdf, grd, idpRange, 'IDW plot of diet.vfishd')
idw_vfishd_2 <- gstat::idw(formula = diet.vfishd~1, locations = perLocation.spdf, newdata = grd, idp=2)

#Jackknife analysis
jackknifeDiffIDP(perLocation.spdf, perLocation.spdf@data$diet.vfishd, 'diet.vfishd', diet.vfishd~1, idpRange)
```


#DIET.VUNK 
TRANSFORMATIONS
```{r}
#Transformation
#examineNormality(diet.vunk)
perLocation.spdf@data$diet.vunk <- diet.vunk

#IDW plots
idpRange = seq(2,4,1)
plotIDW(diet.vunk~1, perLocation.spdf, grd, idpRange, 'IDW plot of diet.vunk')

#Jackknife analysis
jackknifeDiffIDP(perLocation.spdf, perLocation.spdf@data$diet.vunk, 'diet.vunk', diet.vunk~1, idpRange)
```


#DIET.FRUIT 
TRANSFORMATIONS
```{r}
#Transformation
#examineNormality(diet.fruit)
perLocation.spdf@data$diet.fruit <- diet.fruit

#IDW plots
idpRange = seq(2,4,1)
plotIDW(diet.fruit~1, perLocation.spdf, grd, idpRange, 'IDW plot of diet.fruit')

#Jackknife analysis
jackknifeDiffIDP(perLocation.spdf, perLocation.spdf@data$diet.fruit, 'diet.fruit', diet.fruit~1, idpRange)

```


#DIET.NECT 
```{r}
#Transformation
#examineNormality(diet.nect)
perLocation.spdf@data$diet.nect <- diet.nect

#IDW plots
idpRange = seq(2,4,1)
plotIDW(diet.nect~1, perLocation.spdf, grd, idpRange, 'IDW plot of diet.nect')

#Jackknife analysis
jackknifeDiffIDP(perLocation.spdf, perLocation.spdf@data$diet.nect, 'diet.nect', diet.nect~1, idpRange)
```


#DIET.SEED 
```{r}
#Transformation
#examineNormality(diet.seed)
perLocation.spdf@data$diet.seed <- diet.seed

#IDW plots
idpRange = seq(2,4,1)
plotIDW(diet.seed~1, perLocation.spdf, grd, idpRange, 'IDW plot of diet.seed')

#Jackknife analysis
jackknifeDiffIDP(perLocation.spdf, perLocation.spdf@data$diet.seed, 'diet.seed', diet.seed~1, idpRange)

```

#DIET.PLANTO TRANSFORMATIONS
```{r}
#Transformation
#examineNormality(diet.planto)
perLocation.spdf@data$diet.planto <- diet.planto

#IDW plots
idpRange = seq(2,4,1)
plotIDW(diet.planto~1, perLocation.spdf, grd, idpRange, 'IDW plot of diet.planto')

#Jackknife analysis
jackknifeDiffIDP(perLocation.spdf, perLocation.spdf@data$diet.planto, 'diet.planto', diet.planto~1, idpRange)

```


###BASEMAPS
```{r}
makeMap <- function(markerVals, idw, markerName, idwName, transformFormula) {
  medLong = median(perLocation.spdf@coords[,1])
  medLat = median(perLocation.spdf@coords[,2])
  
  markerVals=sort(markerVals)
  lowerQuarter = markerVals[floor(length(markerVals)/4)]
  upperQuarter = markerVals[length(markerVals)-floor(length(markerVals)/4)]
  
  getColor <- function(markerVals) {
    sapply(markerVals, function(currVal) {
    if(currVal <= lowerQuarter) {
      "#FFFFCC"
    } else if(currVal <= upperQuarter) {
      "#41B6C4"
    } else {
      "#0C2C84"
    } })
  }
  
  icons <- awesomeIcons(
    icon = 'ios-close',
    iconColor = getColor(markerVals),
    library = 'ion',
    markerColor = 'white'
  )
  
  # get idw to overlay
  r = raster(idw)
  crs(r) <- sp::CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")
  pal <- colorNumeric(c("#FFFFCC", "#41B6C4", "#0C2C84"), values(r),
    na.color = "transparent")
  
  # draw map
  m = leaflet(perLocation.spdf) %>% setView(lng = medLong, lat = medLat, zoom = 10)
  m %>% addTiles() %>% addAwesomeMarkers(icon=icons, label = as.character(markerName), popup=~as.character(site)) %>%
    addRasterImage(r, colors = pal, opacity = 0.8) %>%
    addLegend(pal = pal, values = values(r), labFormat = labelFormat(transform = transformFormula),
      title = idwName)
}


```

# Make Basemaps
```{r}
#makeMap(perLocation.spdf@data$transformedAbundance, idw_abundance_2, perLocation.spdf@data$abundance, 'abundance',function(x) x^2)
makeMap(perLocation.spdf@data$diet.vfishd, idw_vfishd_2, perLocation.spdf@data$diet.vfishd, 'diet.vfishd',function(x) x)

```

### Moran's I Analyses
```{r}
centroids <- coordinates(perLocation.spdf)
delauney_Scot <- tri2nb(centroids)
summary(delauney_Scot)
SOI_Scot<-graph2nb(soi.graph(delauney_Scot,centroids))
summary(SOI_Scot)

contig_listw <- nb2listw(SOI_Scot, style="B", zero.policy=TRUE) 
moran.test(perLocation.spdf@data$abundance, listw = contig_listw, zero.policy = TRUE)
moran.test(perLocation.spdf@data$speciesRichness, listw = contig_listw, zero.policy = TRUE)
```
